---
title: "Correlation comparison"
author: "Byron Jaeger"
date: "March 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

suppressPackageStartupMessages({
  library(tidyverse)
  library(kableExtra)
  library(geepack)
})

files <- list.files(
  path = 'Code/KableWon',
  all.files = TRUE,
  full.names = TRUE,
  pattern = '.R'
)

for(f in files) source(f)

files <- list.files(
  path='Code/abp_sampling_functions',
  all.files=TRUE, 
  full.names=TRUE, 
  pattern='.R'
)

for(f in files) source(f)

```

## This document

The purpose of this document is to illustrate the method I have applied to compare the associations between left ventricular hypertropy and nocturnal hypertension when (1) a full set of $\geq 5$ blood pressure measurements are used to estimate mean blood pressure during sleep, and (2) a set of $<$ 5 blood pressure measurements are used to estimate mean blood pressure during sleep.

### Part 1

Set a couple of control variables for poisson regression models: 

```{r cars}

control <- c(
  'age',
  'sex',
  'edu',
  'diabetes',
  'currentsmoker',
  'bpmeds',
  'slp_duration'
)

```

Import a dataset with information on left-ventricular hypertrophy status and mean blood pressure during sleep according to a full set of $\geq 5$ blood pressure measurements. 

```{r}

analysis <- file.path(
  "Datasets", "JHS_analysis.RDS"
) %>% 
  readRDS() %>% 
  dplyr::select(
    subjid, # subject id, 
    lvm,    # left-ventricular mass indexed to height
    lvh,    # left-ventricular hypertrophy
    acr,    # albumin-to-creatinine ratio
    albuminuria, # acr > 30
    slp_sbp,# true sleep sbp
    slp_dbp,# true sleep dbp
    !!control # control variables (see above chunk)
  )

print(analysis)

```


Import a dataset that contains blood pressure measurements taken at 1, 2, 3, and 4 AM from participants in the Jackson Heart Study. 

```{r}

model_df <- file.path(
  "Results",
  "Full Results",
  "JHS_FALSE_overall_tsm_1_2_3_4.RDS"
) %>% 
  readRDS() %>% 
  dplyr::select(
    subjid, # subject id
    tsm,    # time since midnight, hours
    sbp,    # systolic blood pressure
    dbp     # diastolic blood pressure
  ) %>% 
  group_by(subjid)

print(model_df)

```


```{r}

model_df <- model_df %>% 
  dplyr::summarize(
    slp_sbp_est=mean(sbp),
    slp_dbp_est=mean(dbp)
  ) %>% 
  # merge these estimates with full analysis data
  left_join(analysis, by = 'subjid') %>% 
  # remove missing values 
  # this is not done in the paper, but
  # it is done here to keep things simple
  na.omit() 

print(model_df)

```

As a quick tangent, note the correlation between estimated nocturnal SBP and SBP and true nocturnal SBP and DBP

```{r}

library(ggcorrplot)

corr_vars <- c(
  "slp_sbp_est",
  "slp_dbp_est",
  "slp_sbp",
  "slp_dbp",
  "lvm",
  "acr"
)

corr <- cor(model_df[,corr_vars])

ggcorrplot(
  corr,
  hc.order = TRUE,
  type = 'lower',
  outline.col = "white",
  lab = TRUE
  ) + 
  theme(
    panel.grid = element_blank()
  )


```


### Part 2

Fit regression models for LVH

```{r}

m1 = model_df %>%
  dplyr::mutate(
    slp_sbp=slp_sbp / 10,
    slp_dbp=slp_dbp / 10
    ) %>% 
  geepack::geeglm(
    formula = lvh ~ slp_sbp + slp_dbp + age + sex + edu + 
      diabetes + currentsmoker + bpmeds + slp_duration,
    family = 'poisson',
    id = 1:nrow(.),
    data = .
  )

m2 = model_df %>%
  dplyr::mutate(
    slp_sbp_est = slp_sbp_est / 10,
    slp_dbp_est = slp_dbp_est / 10
    ) %>% 
  geepack::geeglm(
    formula = lvh ~ slp_sbp_est + slp_dbp_est + age + sex + edu + 
      diabetes + currentsmoker + bpmeds + slp_duration,
    family = 'poisson',
    id = 1:nrow(.),
    data = .
  )

estimates <- list(
  m1, m2
) %>% 
  map(
    broom::tidy
  ) %>% 
  bind_rows(
    .id='model'
  ) %>% 
  dplyr::filter(
    grepl(pattern='slp_sbp', x=term)
  ) %>% 
  mutate(
    pr_est = exp(estimate),
    pr_lwr = exp(estimate - 1.96*std.error),
    pr_upr = exp(estimate + 1.96*std.error)
  )

print(estimates)

```

The effects appear to be almost identical. Here is the same analysis, using albuminuria

```{r}

# model 1 includes all variables except for nht_est
# i.e., model 1 estimates PR for true nht
m1 = model_df %>%
  dplyr::mutate(
    slp_sbp=slp_sbp / 10,
    slp_dbp=slp_dbp / 10
  ) %>% 
  geepack::geeglm(
    formula = albuminuria ~ slp_sbp + slp_dbp + age + sex + edu + 
      diabetes + currentsmoker + bpmeds + slp_duration,
    family = 'poisson',
    id = 1:nrow(.),
    data = .
  )

# model 2 includes all variables except for nht
# i.e., model 1 estimates PR for estimated nht
m2 = model_df %>%
  dplyr::mutate(
    slp_sbp_est = slp_sbp_est / 10,
    slp_dbp_est = slp_dbp_est / 10
  ) %>% 
  geepack::geeglm(
    formula = albuminuria ~ slp_sbp_est + slp_dbp_est + age + sex + edu + 
      diabetes + currentsmoker + bpmeds + slp_duration,
    family = 'poisson',
    id = 1:nrow(.),
    data = .
  )

estimates <- list(
  m1, m2
) %>% 
  map(
    broom::tidy
  ) %>% 
  bind_rows(
    .id='model'
  ) %>% 
  dplyr::filter(
    grepl(pattern='slp_sbp', x=term)
  ) %>% 
  mutate(
    pr_est = exp(estimate),
    pr_lwr = exp(estimate - 1.96*std.error),
    pr_upr = exp(estimate + 1.96*std.error)
  )

print(estimates)

```

## Part 3

Bootstrap the difference in prevalence ratios. The code is omitted because it is just a loop of the code above.

```{r, echo = FALSE}


estimate_difference <- function(data, outcome, compare, control)
{
  
  # n <- nrow(data)
  # boot_indx <- sample(n, replace=TRUE)
  
  model_formulas <- map(
    compare, .f=function(.variable){
      as.formula(
        paste(
          outcome, '~', .variable, '+', paste(control, collapse=' + ')
        ) 
      )
    }
  )
  
  model_estimates <- map2_dfr(
    model_formulas, compare, .f=function(.formula,.variable){
      geepack::geeglm(
        formula = .formula,
        family = 'poisson',
        id = 1:nrow(data),
        data = data
      ) %>% 
        broom::tidy() %>% 
        dplyr::filter(term==.variable)
    }
  )
  
  model_estimates %>% 
    pluck("estimate") %>% 
    diff()
}

set.seed(32989)
library(bcaboot)

results <- c("lvh","albuminuria") %>% 
  set_names(paste(.)) %>% 
  map(
    .f=function(.outcome){
      bcajack(
        x = model_df, 
        B = nrow(model_df), 
        func = estimate_difference, 
        verbose = FALSE,
        compare = c("slp_sbp","slp_sbp_est"),
        outcome = .outcome,
        control = control
      )
    }
  )

```

```{r}

results$lvh$lims %>% 
  as_tibble(rownames = 'bca_pct') %>% 
  dplyr::select(
    BCA = bca, 
    `BCA percentile` = bca_pct,
    standard = std,
    `standard percentile` = pct
  ) %>% 
  kable(
    caption = 'Bias corrected accelerated (BCA) and standard bootstrap distributions for the difference in association with left-ventricular hypertrophy between estimated and true nocturnal systolic blood pressure.',
    digits=3,
    align = 'c'
  ) %>% 
  kable_styling()

```

The first column shows the estimated bias-corrected accelerated (BCA) confidence limits at the requested alpha percentiles which can be compared with the standard limits $\theta \pm \hat{\sigma} \cdot z_\alpha$ under the column titled standard.  The percentile column gives percentiles of the ordered `r n_boots` bootstrap replications corresponding to the Bca limits, e.g. the 0.502 percentile equals the .500 Bca limit: -0.713.

```{r}

results$albuminuria$lims %>% 
  as_tibble(rownames = 'bca_pct') %>% 
  dplyr::select(
    BCA = bca, 
    `BCA percentile` = bca_pct,
    standard = std,
    `standard percentile` = pct
  ) %>% 
  kable(
    caption = 'Bias corrected accelerated (BCA) and standard bootstrap distributions for the difference in association with albuminuria between estimated and true nocturnal systolic blood pressure.',
    digits=3,
    align = 'c'
  ) %>% 
  kable_styling()


```

